---
title: "epimutacionsData: a  data repostory for epimutacions package"
author: "Leire Abarrategui"
date: "`r doc_date()`"
vignette: >
  %\VignetteIndexEntry{References for epimutacions using Bioconductor's ExperimentHub}
  %\VignetteEngine{knitr::rmarkdown}
output: 
  BiocStyle::html_document

bibliography: refs.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Overview

The `epimutacionsData` package is a repository of  datasets for the `epimutacions` package. It includes 2 datasets to use as an example: (1) Reference panel and (2) case and control samples. Also, a dataset specifying all the candidate epimutations is included. 

## Candidate epimutations

In Illumina 450K array [@reproducibilityinfinium], probes are unequally distributed along the genome, limiting the number of regions that can fulfil the requirements to be considered an epimutation. So, we have computed a dataset containing the regions that are candidates to become an epimutation. 

To define the candidate epimutations, we relied on the clustering from bumphunter [@jaffe2012bump]. We defined a primary dataset with all the CpGs from the Illumina 450K array. Then, we run bumphunter and selected those regions with at least 3 CpGs. As a result, we found 40408 candidate epimutations which are available in the `candRegsGR` dataset.

## Example datasets

### Reference panel

The package includes an `RGChannelSet` class reference panel (`reference_panel`) which contains 22 whole cord blood samples from healthy children born via caesarian from the GSE127824 cohort [@gervin2019systematic].

### Control and case samples

The `methy` dataset includes 51 DNA methylation profiling of whole blood samples. 48 controls from GSE104812 [@shi2018dna] cohort and 3 cases from GSE97362 [@butcher2017charge]. it is a `GenomicRatioSet` class object. 


# sessionInfo()

```{r sessionInfo}
sessionInfo()
```

# References